---
title: "Canidae tests"
author: "D. Rex Mitchell"
date: "2023-02-14"
output: html_document
---

################################################################################
Import packages
```{r}
library(geomorph)
library(landvR)
library(ape)
library(phytools)
```

Code for color scheme function
```{r}
map2color <- function(x, pal, limits = range(x)){
  pal[findInterval(x, seq(limits[1], limits[2], length.out = length(pal) + 1), 
                   all.inside=TRUE)]
}
```

Import chosen taxon .rda file
```{r}
load("Canidae.rda")
```

Perform Procrustes superimposition of shape data, extract shape
```{r}
gpa <- gpagen(data)
shape <- gpa$coords
```

Obtain centroid size scaling index for the taxon (largest/smallest)
```{r}
size_scale <- range(size)[2]/range(size)[1]; size_scale
```
Plot tree with size
```{r}
obj <- contMap(tree, size, legend=FALSE)
plot(obj)
map2 <- setMap(obj, c("blue", "green","yellow","orange","red"))
plot(map2, mar=c(5.1,0.2,0.2,0.2), legend=20, leg.txt="cranial centroid size")
```

#### Vulpes branch ###
```{r}
datavul <- data[,,c(18:19,21:30)] # n = 12
dimnames(datavul)[[3]]
gpavul <- gpagen(datavul)
sizevul <- size[c(18:19,21:30)]
```

```{r}
size_scale <- range(sizevul)[2]/range(sizevul)[1]; size_scale # 15.37666 - 29.04415 = 1.889
```

Test for allometry in Vulpes clade.
```{r}
allom <- procD.lm(gpavul$coords~log(sizevul), iter=999); summary(allom)
allom_plot <- plot(allom, type = "regression", predictor = log(sizevul), 
                   reg.type = "RegScore", pch = 22, bg="grey", 
                   xlab = "log(centroid size)", ylab = "Regression score", cex=2)
### Allometry significant (Rsq=0.220, F=2.814, p=0.044)
```
Predict allometric shape variation for the Vulpes clade
```{r}
preds <- shape.predictor(allom$GM$fitted, x = allom_plot$PredLine, 
                         predmin = min(allom_plot$PredLine), 
                         predmax = max(allom_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmax, preds$predmin, col = heat.colors, col.val = differences[[1]][1:30], magnitude = 1, pt.size = 3)
### larger species have a more narrow cranium and slight premaxillae projection
```

# Canis branch
```{r}
datacan <- data[,,c(2:10,20,17)] # n = 11
gpacan <- gpagen(datacan)
sizecan <- size[c(2:10,20,17)]
```

```{r}
size_scale <- range(sizecan)[2]/range(sizecan)[1]; size_scale # 1.755
```
Test for allometry in the Canis clade
```{r}
allom <- procD.lm(gpacan$coords~log(sizecan), iter=999); summary(allom)
### Allometry not significant (Rsq=0.159, F=1.696, p=0.16)
```

Plot PCA to explore relationships
```{r}
PCA <- gm.prcomp(gpacan$coords)
cont <- colorRampPalette(c("hotpink","darkblue"), interpolate = "linear")
colage <- cont(15)
colmap <- map2color(x=size, pal=colage, limits=range(size))
plot(PCA$x[,2]~PCA$x[,1], pch=21, cex=size/20, bg=colmap)
text(PCA$x[,1],PCA$x[,2], labels = dimnames(gpacan$coords)[[3]])
# PC1 defines facial gracility and separates small prey specialists from large prey specialists.Canis simensis, a small prey speicialist, has the leanest  cranium.
```
Examine PC1 shape variation. 
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp1$max,
                                      reference = PCA$shapes$shapes.comp1$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp1$min, PCA$shapes$shapes.comp1$max, col = heat.colors, col.val = differences[[1]][1:dim(gpacan$coords)[1]], magnitude = 1, pt.size = 2)
# PC1 represents cranial width
```

Vulpes branch + large prey specialists
```{r}
datavul2 <- data[,,c(9,17:20,21:30)] # n = 15
gpa <- gpagen(datavul2)
sizevul2 <- size[c(9,17:20,21:30)]
```

```{r}
size_scale <- range(sizevul2)[2]/range(sizevul2)[1]; size_scale # 2.308
```

```{r}
allom <- procD.lm(gpa$coords~log(sizevul2), iter=999); summary(allom)
allom_plot <- plot(allom, type = "regression", predictor = log(sizevul2), 
                   reg.type = "RegScore", pch = 22, bg="grey", 
                   xlab = "log(centroid size)", ylab = "Regression score", cex=2)
### Allometry significant (Rsq = 0.175, F=2.751, p=0.044)
```

# Allometry predict shape
```{r}
preds <- shape.predictor(allom$GM$fitted, x = allom_plot$PredLine, 
                         predmin = min(allom_plot$PredLine), 
                         predmax = max(allom_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmax, preds$predmin, col = heat.colors, col.val = differences[[1]][1:30], magnitude = 1, pt.size = 2)
# Allometric prediction for more stout crania with increased size
```

